---
title: "Homework 2"
author: Rachel Tao
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

# Problem 1

```{r}
trashwheel_df <- 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data!

```{r}
precip_2018 <- 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 <- 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df <- 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df <- 
  bind_rows(precip_2018, precip_2017) 

month_df <- 
  left_join(precip_df, month_df, by = "month") %>% 
  select(-"month")

```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include monthly precipitation data. The total precipitation in 2018 was `r   filter(month_df, year == "2018") %>% pull(total) %>% sum()`. The median number of sports balls in a dumpster in 2017 was `r filter(trashwheel_df, year == 2017) %>% pull(sports_balls) %>% median()`.

# Problem 2

Read and clean NYC subway data.

```{r}
nycsubway_df <- 
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  mutate(
    entry = 
      ifelse(entry == "YES", TRUE, FALSE),
    vending = 
      ifelse(vending == "YES", TRUE, FALSE))
```

This dataset has `r ncol(nycsubway_df)` columns and `r nrow(nycsubway_df)` rows, and includes variables line, station name, station latitude, station longitude, routes served (route1-route11), entry, entrance type, vending, and ADA compliance. For data cleaning, I used the function clean_names in the janitor package to make the column names more readable in R, I selected only the columns for variables that were relevant for this problem, and I used the mutate function and ifelse function to change the entry column from character to logical. Since a future question asks about vending, I used the same method to change the vending column from character to logical. These data are cleaner than before, but they are not tidy yet. They would be tidier if we could condense the routes so that we did not have 11 columns for route1 through route11, and instead could have one column for route number (1-11), and another for route name (A, N, R, etc.).

There are `r nrow(distinct(nycsubway_df, line, station_name))` distinct stations, of which `r distinct(nycsubway_df, line, station_name, .keep_all = TRUE) %>% filter(ada == TRUE) %>% nrow()` are ADA compliant. The proportion of station entrances/exits without vending that allow entrance is `r (filter(nycsubway_df, vending == FALSE) %>% filter(entry == TRUE) %>% nrow())/(filter(nycsubway_df, vending == FALSE) %>% nrow())`.

Reformat data so that route number and route name are distinct variables:

```{r}

nycsubway_df <- 
  mutate(nycsubway_df,
         route8 = 
           as.character(route8),
         route9 =
           as.character(route9),
         route10 = 
           as.character(route10),
         route11 = 
           as.character(route11)) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )

```

`r distinct(nycsubway_df, line, station_name, .keep_all = TRUE) %>% filter(route_name == 'A') %>% nrow()` distinct stations serve the A train. Of stations that serve the A train, `r distinct(nycsubway_df, line, station_name, .keep_all = TRUE) %>% filter(route_name == 'A') %>% filter(ada == TRUE) %>% nrow()` are ADA compliant.

# Problem 3

Read in and clean pols-month.csv

```{r}

pols_df <- read_csv(
  "./data/fivethirtyeight_datasets/pols-month.csv") %>% 
    separate(mon, into = c("year", "month", "day"), sep = ) %>% 
  mutate(
    month = 
      as.integer(month)
  ) %>% 
  relocate(year, month, prez_gop, prez_dem) %>% 
  pivot_longer(
    prez_gop:prez_dem,
    names_to = "president",
    names_prefix = "prez_",
    values_to = "prez_value") %>%
  filter(prez_value == 1) %>% 
  select(-day, -prez_value)

month_df <- 
  tibble(
    month = 01:12,
    month_name = month.name
  )

pols_df <- 
  left_join(pols_df, month_df, by = "month") %>% 
  select(-month) %>% 
  rename(month = month_name) %>% 
  mutate(
    year =
      as.double(year)
  ) %>% 
  relocate(year, month)

```

Read in and clean snp.csv

```{r}
snp_df <- read_csv(
  "./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, into = c("month", "day", "year"), sep = ) %>% 
  mutate(
      month =
        as.integer(month)
  ) %>% 
  select(-day)

month_df <- 
  tibble(
    month = 1:12,
    month_name = month.name
  )

snp_df <- 
  left_join(snp_df, month_df, by = "month") %>% 
  select(-month) %>% 
  rename(month = month_name) %>% 
  mutate(
    year = 
      as.double(year)
  ) %>% 
  relocate(year, month)

```

Read in and clean unemployment.csv

```{r}
unemployment_df <- read_csv(
  "./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
  )

month_df <- 
  tibble(
    month = month.abb,
    month_name = month.name
  )

unemployment_df <- 
  left_join(unemployment_df, month_df, by = "month") %>% 
  select(-month) %>% 
  rename(month = month_name) %>% 
  rename(year = Year) %>% 
  relocate(year, month)

```

Join datasets snp, pols, and unemployment.

```{r}

fivethirtyeight_df <- 
  left_join(pols_df, snp_df, by = c("year", "month")) %>% 
  left_join(unemployment_df, by = c("year", "month"))

```

The pols_df dataset included information on how many governors, senators, and congressional representatives were democrats or republicans at any given time from `r select(pols_df, year) %>% min()` to `r select(pols_df, year) %>% max()`. The snp_df dataset included closing values of the S&P stock index from `r select(snp_df, year) %>% min()` to `r select(snp_df, year) %>% max()` for specific dates of observation during that period. The unemployment_df dataset included monthly percentage of unemployment from `r select(unemployment_df, year) %>% min()` to `r select(unemployment_df, year) %>% max()`. The fivethirtyeight_df dataset includes information from all three datasets, and has `r ncol(fivethirtyeight_df)` columns and `r nrow(fivethirtyeight_df)` rows. It includes data from `r select(fivethirtyeight_df, year) %>% min()` to `r select(fivethirtyeight_df, year) %>% max()`, although some columns have missing values for certain years, because the original datasets did not have fully-overlapping time frames. The variables in this dataset include: year, month, number of republican governors ('gov_gop'), number of republican senators ('sen_gop'), number of republican congressional representatives ('rep_gop'), number of democratic governors ('gov_dem'), number of democratic senators ('sen_dem'), number of democratic congressional representatives ('rep_dem'), closing value of S&P stock index ('close'), and monthly percentage unemployment ('unemployment'). The variable 'close' in particular includes the closing value of the S&P stock index on a particular date within the month it is associated with, as opposed to an average closing value for the month or some other value.
