---
title: "Homework 2"
author: Rachel Tao
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

# Problem 1

```{r}
trashwheel_df <- 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data!

```{r}
precip_2018 <- 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 <- 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df <- 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df <- 
  bind_rows(precip_2018, precip_2017) 

month_df <- 
  left_join(precip_df, month_df, by = "month") %>% 
  select(-"month")

```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include monthly precipitation data. The total precipitation in 2018 was `r   filter(month_df, year == "2018") %>% pull(total) %>% sum()`. The median number of sports balls in a dumpster in 2017 was `r filter(trashwheel_df, year == 2017) %>% pull(sports_balls) %>% median()`.

# Problem 2

Read and clean NYC subway data.

```{r}
nycsubway_df <- 
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  mutate(
    entry = 
      ifelse(entry == "YES", TRUE, FALSE),
    vending = 
      ifelse(vending == "YES", TRUE, FALSE))
```

This dataset has `r ncol(nycsubway_df)` columns and `r nrow(nycsubway_df)` rows, and includes variables line, station name, station latitude, station longitude, routes served (route1-route11), entry, entrance type, vending, and ADA compliance. For data cleaning, I used the function clean_names in the janitor package to make the column names more readable in R, I selected only the columns for variables that were relevant for this problem, and I used the mutate function and ifelse function to change the entry column from character to logical. Since a future question asks about vending, I used the same method to change the vending column from character to logical. These data are tidy.

```{r}
ada_stations <- 
  distinct(nycsubway_df, line, station_name, .keep_all = TRUE) %>% 
  filter(ada == TRUE)

no_vending <- 
  filter(nycsubway_df, vending == FALSE) %>% 
  nrow()

entry_no_vending <- 
  filter(nycsubway_df, vending == FALSE) %>% 
  filter(entry == TRUE) %>% 
  nrow()
```

There are `r nrow(distinct(nycsubway_df, line, station_name))` distinct stations, of which `r nrow(ada_stations)` are ADA compliant. The proportion of station entrances/exits without vending that allow entrance is `r entry_no_vending/no_vending`.